Neighbor-Based Decentralized Training Strategies for Multi-Agent Reinforcement Learning

### 2024-10-12 paper-2024-sac-marl-neighbouring-strategies.tex
shrink: dtde e ctde images
shrink: discussion on independent learners

\subsection{Learning Schemes}
\begin{figure}
  \centering
  \includegraphics[width=0.7\linewidth]{figures/CTDE.pdf}
  \vskip 1 mm
  \includegraphics[width=0.7\linewidth]{figures/DTDE.pdf}
  \caption{Graphical representation of centralized training with decentralized execution (left) and decentralized training with decentralized execution (rigth) in multi-agent reinforcement learning.}
  \label{fig:learning_paradigms}
\end{figure}
Multi-agent learning can be categorized based on the training scheme and the information accessibility for agents. 
Two prominent paradigms are centralized training with decentralized execution (CTDE) and decentralized training with decentralized execution (DTDE)---see \Cref{fig:learning_paradigms}.

@@

\subsection{State of the Art}\label{sec:sota}
MARL is a longstanding research area with a rich literature, starting 
from seminal works like~\cite{tan1993multi} and~\cite{busoniu2008comprehensive} to more recent contributions like~\cite{gronauer2022multi} and~\cite{canese2021multi}.
%
In the following, we briefly discuss some of the most relevant approaches in the field, which lead to the motivation for our work.
\subsubsection{Independent Learners}
In independent learners, each agent learns its policy independently, disregarding the actions and policies of other agents. 
This straightforward approach, among the earliest explored in MARL, often relies on variants of Q-learning. 
%
Several algorithms have attempted to address these coordination challenges within the independent learner paradigm. 
Basic decentralized Q-learning~\cite{tan1993multi} allows each agent to learn its Q-values independently but often converges to suboptimal joint policies due to a lack of coordination. 
Subsequent work introduced heuristic-based extensions to improve coordination.
Approaches like distributed Q-learning~\cite{lauer2000algorithm} incorporate optimism in Q-value updates and equilibrium selection mechanisms to mitigate the convergence to shadowed equilibria (i.e., equilibria that are unstable due to the independent learning processes). 
Instead, algorithms such as hysteretic Q-learning~\cite{matignon2007hysteretic}, lenient Q-Learning~\cite{bloembergen2010lenient} and win-or-learn fast policy hill climbing~\cite{bowling2002multiagent} employ adaptive learning rates to address the issue of alternating exploration between agents, improving robustness and convergence.
%
Empirical comparisons of these algorithms highlight the trade-offs in performance across various MARL tasks and coordination challenges~\cite{matignon2012independent}. 
While initially developed for tabular settings, some of these methods, like hysteretic Q-learning~\cite{palmer2017lenient} and lenient Q-learning, have been extended to deep reinforcement learning settings. 
However, these independent approaches, even with deep learning, often exhibit instability and slow convergence in complex environments and fundamentally lack explicit mechanisms for inter-agent coordination. 
%Consequently, more sophisticated MARL algorithms have been developed that explicitly account for the interactions and policies of other agents to facilitate more effective coordination.
\subsubsection{Neighboring-Based Methods}
Neighboring-based methods in MARL address the scalability and communication overhead challenges of fully centralized approaches by restricting agent interactions to a local neighborhood. 
This introduces agents communication and coordination with a limited set of neighbors, reducing the complexity of full joint action and policy learning.
%
One prominent category within neighboring-based methods is \emph{consensus-based} algorithms. 
These algorithms aim to reach a consensus among agents on shared values or policies through local information exchange. 
Early work like QD-learning~\cite{kar2012qd} extends Q-learning with a consensus update term, 
enabling agents to converge towards optimal Q-values despite only receiving local rewards.
%  
Recent advances in consensus-based MARL leverage deep learning and actor-critic architectures, 
combining local policy updates with consensus updates on critic parameters or gradients, 
as seen in the recent works~\cite{7434032,pennesi2010distributed,zhang2018fully,DBLP:journals/jmlr/CiosekW20}.

>>>>

\subsection{Learning Schemes}
\begin{figure}
  \centering
  \includegraphics[width=0.49\linewidth]{figures/CTDE.pdf}
  %\vskip 1 mm
  \includegraphics[width=0.49\linewidth]{figures/DTDE.pdf}
  \caption{Graphical representation of CTDE (left) and DTDE (right) in MARL.}
  \label{fig:learning_paradigms}
\end{figure}
Multi-agent learning can be categorized based on the training scheme and the information accessibility for agents. 
Two prominent paradigms are centralized training with decentralized execution (CTDE) and decentralized training with decentralized execution (DTDE)---see \Cref{fig:learning_paradigms}.

@@

\subsection{State of the Art}\label{sec:sota}
MARL is a longstanding research area with a rich literature, starting 
from seminal works like~\cite{tan1993multi} and~\cite{busoniu2008comprehensive} to more recent contributions like~\cite{gronauer2022multi} and~\cite{canese2021multi}.
%
In the following, we briefly discuss some of the most relevant approaches in the field, which lead to the motivation for our work.
\subsubsection{Independent Learners}
In independent learners, each agent learns its policy independently, disregarding the actions and policies of other agents. 
This straightforward approach, among the earliest explored in MARL, often relies on variants of Q-learning. 
%
Several algorithms have attempted to address these coordination challenges within the independent learner paradigm. 
Basic decentralized Q-learning~\cite{tan1993multi} allows each agent to learn its Q-values independently but often converges to suboptimal joint policies due to a lack of coordination. 
%Subsequent work introduced heuristic-based extensions to improve coordination.
%Approaches like distributed Q-learning~\cite{lauer2000algorithm} incorporate optimism in Q-value updates and equilibrium selection mechanisms to mitigate the convergence to shadowed equilibria (i.e., equilibria that are unstable due to the independent learning processes). 
%Instead, algorithms such as hysteretic Q-learning~\cite{matignon2007hysteretic}, lenient Q-Learning~\cite{bloembergen2010lenient} and win-or-learn fast policy hill climbing~\cite{bowling2002multiagent} employ adaptive learning rates to address the issue of alternating exploration between agents, improving robustness and convergence.
%
%Empirical comparisons of these algorithms highlight the trade-offs in performance across various MARL tasks and coordination challenges~\cite{matignon2012independent}. 
Subsequent work introduced heuristic-based extensions to improve coordination, like distributed Q-learning~\cite{lauer2000algorithm}, hysteretic Q-learning~\cite{matignon2007hysteretic}, lenient Q-Learning~\cite{bloembergen2010lenient}, however empirical comparisons of these algorithms highlight the trade-offs in performance across various MARL tasks and coordination challenges~\cite{matignon2012independent}. 
While initially developed for tabular settings, some of these methods, like hysteretic Q-learning~\cite{palmer2017lenient} and lenient Q-learning, have been extended to deep reinforcement learning settings. 
However, these independent approaches, even with deep learning, often exhibit instability and slow convergence in complex environments and fundamentally lack explicit mechanisms for inter-agent coordination. 
%Consequently, more sophisticated MARL algorithms have been developed that explicitly account for the interactions and policies of other agents to facilitate more effective coordination.
\subsubsection{Neighboring-Based Methods}
Neighboring-based methods in MARL address the scalability and communication overhead challenges of fully centralized approaches by restricting agent interactions to a local neighborhood. 
This introduces agents communication and coordination with a limited set of neighbors, reducing the complexity of full joint action and policy learning.
%
One prominent category within neighboring-based methods is \emph{consensus-based} algorithms. 
These algorithms aim to reach a consensus among agents on shared values or policies through local information exchange. 
Early work like QD-learning~\cite{kar2012qd} extends Q-learning with a consensus update term, 
enabling agents to converge towards optimal Q-values despite only receiving local rewards.
%  
Recent advances in consensus-based MARL leverage deep learning and actor-critic architectures, 
combining local policy updates with consensus updates on critic parameters or gradients, 
as seen in the recent works~\cite{7434032,pennesi2010distributed,zhang2018fully,DBLP:journals/jmlr/CiosekW20}.
